---
title: "Extracting diseases from Electronic Health Records"
author: "Eric Wanjau"
toc: true
format: 
  html:
    number-sections: true
    toc-location: left
    code-tools: true
    code-fold: false
    code-link: true
editor: visual
execute: 
  warning: false
  message: false
bibliography: references.bib
---

This notebook explores how MedCAT can be used for extracting concepts from unstructured texts and follows/modifies the tutorials at: [`MedCAT tutorials`](https://github.com/CogStack/MedCATtutorials)

MedCAT is an open source annotation tool that can learn to extract concepts (e.g. disease, symptoms, medications) from free-text and link them to any biomedical ontology such as SNOMED-CT and UMLS [@kraljevic2021].

::: callout-note
This notebook uses both R and Python at different sections.
Please see:

<https://rpubs.com/eR_ic/transfoRmers> on how to set up reticulate [@reticulate] and Python environments.
:::

```{r echo=FALSE, include=FALSE}
# Load reticulate into current R session
library(reticulate)
library(here)
use_python("env/Scripts/python.exe")

# Retrieve/force initialization of Python
reticulate::py_config()
```

```{r eval=FALSE}
# Load reticulate into current R session
library(reticulate)
library(here)
use_python("env/Scripts/python.exe")

# Retrieve/force initialization of Python
reticulate::py_config()

```

```{r}
# Check if python is available
reticulate::py_available()
```

With everything set up, let's begin extracting concepts from unstructured texts.

```{python}
#| eval: false

# Install medcat into virtual environment
!pip install medcat

```
